---
title: "DSII_final_data"
author: "Carolina Downie"
date: "4/21/2018"
output: html_document
---

```{r, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(janitor)
library(stringr)
library(ggridges)
library(ggthemes)
library(forcats)
library(viridis)
library(plotly)
library(knitr)
library(kableExtra)
library(shiny)
library(rvest)
library(httr)
```


```{r loading hospital_2 dataset}
# Load hospital_data_2.csv below for results of above code chunk
hospital_data_2 <- read_csv("data/hospital_data_2.csv")
```

#DON'T NEED TO RE-RUN THIS CHUNK, HOSPITAL_DATA_REGIONS FILE HAS BEEN CREATED
```{r assigning states to region categories, writing to new file}

hospital_data_regions <- hospital_data_2 %>%
  mutate(region = ifelse(grepl("ME|NH|VT|MA|RI|CT", state), "New England", ifelse(grepl("NY|PA|NJ", state), "Middle Atlantic", ifelse(grepl("WI|MI|IL|IN|OH", state), "East North Central", ifelse(grepl("ND|SD|NE|KS|MN|IA|MO", state), "West North Central", ifelse(grepl("DE|MD|DC|VA|WV|NC|SC|GA|FL", state), "South Atlantic", ifelse(grepl("KY|TN|MS|AL", state), "East South Central", ifelse(grepl("OK|TX|AR|LA", state), "West South Central", ifelse(grepl("ID|MT|WY|NV|UT|CO|AZ|NM", state), "Mountain", ifelse(grepl("AK|WA|OR|CA|HI", state), "Pacific", "U.S. Territories")))))))))) 
  

write.csv(hospital_data_regions, "./data/hospital_data_regions.csv")

```


Loading hospital_data_regions, adding numerical coding of geographic region, hospital type, hospital_ownership, patient experience, and hospital overall rating
```{r loading hospital data regions, recoding}

hospital_data_regions <- read_csv("data/hospital_data_regions.csv")

#Recoding hospital_data_regions so categorical data is numeric/factor instead of text
hospital_data_regions_numeric <- hospital_data_regions %>% mutate(region_num = ifelse(grepl("New England", region), "1", ifelse(grepl("Middle Atlantic", region), "2", ifelse(grepl("East North Central", region), "3", ifelse(grepl("West North Central", region), "4", ifelse(grepl("South Atlantic", region), "5", ifelse(grepl("East South Central",region), "6", ifelse(grepl("West South Central", region), "7", ifelse(grepl("Mountain", region), "8",ifelse(grepl("Pacific", region), "9", ifelse(grepl("U.S. Territories", region), "10", "0"))))))))))) %>% 
  mutate(hospital_type_num = ifelse(grepl("Critical Access Hospitals", hospital_type), "1", "2")) %>%
  mutate(hospital_ownership_num = ifelse(grepl("Government - Hospital District or Authority", hospital_ownership), "1", ifelse(grepl("Proprietary", hospital_ownership), "2", ifelse(grepl("Voluntary non-profit - Other", hospital_ownership), "3", ifelse(grepl("Voluntary non-profit - Private", hospital_ownership), "4", ifelse(grepl("Voluntary non-profit - Church", hospital_ownership), "5", ifelse(grepl("Government - Local", hospital_ownership), "6", ifelse(grepl("Government - State", hospital_ownership), "7", ifelse(grepl("Physician", hospital_ownership), "8", ifelse(grepl("Government - Federal", hospital_ownership), "9", ifelse(grepl("Tribal", hospital_ownership), "10", "0"))))))))))) %>%
  mutate(patient_experience_num = ifelse(grepl("Not Available", patient_experience_national_comparison), "1", ifelse(grepl("Below the national average", patient_experience_national_comparison), "2", ifelse(grepl("Same as the national average", patient_experience_national_comparison), "3", ifelse(grepl("Above the national average", patient_experience_national_comparison), "4", "0"))))) %>% 
  mutate(region_num = as.factor(region_num), hospital_type_num = as.factor(hospital_type_num), hospital_ownership_num = as.factor(hospital_ownership_num), patient_experience_num = as.factor(patient_experience_num))

```
 
 

#UNSUPERVISED LEARNING AND EXPLORATORY DATA ANALYSIS

Exploratory data analysis ...

*About the dataset*

The main dataset that we chose to analyze was the [Hospital General Information](https://data.medicare.gov/Hospital-Compare/Hospital-General-Information/xubh-q36u/data) dataset, which is published by the Centers for Medicare & Medicaid Services and contains information about all hospitals in the United States (and U.S. Territories) that have been registered with Medicare. According to the Medicare [website](https://data.medicare.gov/Hospital-Compare/Hospital-General-Information/xubh-q36u/about), this dataset was created in May 2014 and was last updated in October 2017. We downloaded the dataset as a csv file from [Data.Medicare.gov](https://data.medicare.gov/Hospital-Compare/Hospital-General-Information/xubh-q36u/data) by clicking on the "Export" tab and selecting the CSV option from the Download section, which allowed us to download a copy of the dataset in a static format. 

[Hospital Compare](https://www.cms.gov/medicare/quality-initiatives-patient-assessment-instruments/hospitalqualityinits/hospitalcompare.html) is a program that was created in 2002 to provides "information on how well hospitals provide recommended care to their patients," based on a variety of measures including patient experiences, readmissions & death rates, payment & value of care. In 2016, "Overall Hospital Quality Star Rating" was added to the Hospital Compare metrics. More information about hospital ratings methodology can be found [here](https://www.qualitynet.org/dcs/ContentServer?c=Page&pagename=QnetPublic%2FPage%2FQnetTier3&cid=1228775957165) and [here](http://www.hcahpsonline.org/Files/October_2017_Star%20Ratings_Tech%20Notes.pdf). 

```{r}
# hospital rating based on hospital type (boxplot)

hospital_data_regions_numeric %>% 
   mutate(hospital_type = fct_reorder(hospital_type, hospital_overall_rating)) %>%
ggplot(aes(x = hospital_type, y = hospital_overall_rating, group = hospital_type)) + geom_boxplot(aes(color = hospital_type)) + labs(title = "Distribution of Hospital Rating Based on Hospital Type", x = "Hospital Type", y = "Hospital Rating") + theme_classic() + theme(axis.text.x = element_text(hjust = 1), plot.title = element_text(hjust = 0.5))


# hospital rating based on ownership (boxplot)
hospital_data_regions_numeric %>% 
   mutate(hospital_ownership = fct_reorder(hospital_ownership, hospital_overall_rating)) %>%
ggplot(aes(x = hospital_ownership, y = hospital_overall_rating, group = hospital_ownership)) + geom_boxplot(aes(color = hospital_ownership)) + labs(title = "Distribution of Hospital Rating Based on Hospital Ownership", x = "Hospital Ownership", y = "Hospital Rating") + theme_classic() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(axis.text.x = element_text(hjust = 1), plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")


# hospital rating distribution (barplot)
hospital_data_regions_numeric %>% 
  mutate(hospital_overall_rating = as.character(hospital_overall_rating)) %>%
  count(hospital_overall_rating) %>% 
  mutate(hospital_overall_rating = fct_reorder(hospital_overall_rating, n)) %>% 
  plot_ly(x = ~hospital_overall_rating, y = ~n, color = ~hospital_overall_rating, type = "bar", colors = "Set2") %>%
   layout(legend = list(x = 100, y = 0.5)) %>%
    layout(
    title = "Distribution of Hospital Rating") %>%
    layout(yaxis = list(title = 'Total number of Hospitals')) %>%
    layout(xaxis = list(title = 'Hospital Rating'))




# hospital rating and Patient Experience Level (Scatter plot)
scatter_patient = hospital_data_2 %>%
  filter(!(patient_experience_national_comparison == "Not Available")) %>%
  ggplot(aes(x = patient_experience_national_comparison, y = hospital_overall_rating, color = hospital_overall_rating)) +
  geom_point(alpha = 0.25) +
  scale_color_viridis() +
  coord_cartesian() + labs(title = "Hospital Rating and Patient Experience level", x = "Patient Experience Level", y = "Hospital Rating") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "none")

ggplotly(scatter_patient)


#Geographic distribution of patient scores

hospital_data_regions_numeric %>% ggplot(aes(x = hospital_overall_rating)) + geom_histogram() + facet_wrap(~region)
```






K-Means Clustering
```{r}
#Plotting region number by hospital ownership number, color by hospital_overall_rating

hospital_data_regions_numeric %>% mutate(hospital_overall_rating = as.factor(hospital_overall_rating)) %>% mutate(hospital_ownership_num = as.factor(hospital_ownership_num)) %>% ggplot(aes(x = region_num, y = hospital_ownership_num, color = hospital_overall_rating)) + geom_point() + labs(title = "Plotting region number by hospital ownership number")

##NOTE: Current plot axes aren't in order. Is this plot even necessary? 


#Now need to apply clustering algorithm
km_hospitals_train <- hospitals_train %>% select(region_num, hospital_type_num, hospital_ownership_num, patient_experience_num, hospital_overall_rating)
km_hospitals <- kmeans(hospitals_train, 5, nstart = 20)
km_hospitals_cluster <- km_hospitals$cluster

#Plotting clusters
plot(km_hospitals_train, col = (km_hospitals$cluster + 1)) + title("K Means clustering")


```



##SUPERVISED LEARNING


#Creating training and test sets--training 70% of data set, test set remaining dataset
```{r}
set.seed(5)
train <- sample(1:nrow(hospital_data_regions_numeric), 2489)

hospitals_train <- hospital_data_regions_numeric[train,]

hospitals_test <- hospital_data_regions_numeric[-train,]

```



#Creating REGRESSION TREE
```{r}
library(tree)

#Creating regression tree
hospital_tree_all <- tree(hospital_overall_rating ~ region_num + hospital_type_num + hospital_ownership_num + patient_experience_num, data = hospitals_train)

plot(hospital_tree_all)
text(hospital_tree_all, pretty = 0)

```

Cross-validation of tree
```{r cross-validation of the tree}
cv_hospital_tree_all <- cv.tree(hospital_tree_all)

plot(cv_hospital_tree_all)
text(cv_hospital_tree_all)

prune_hospital_train_all <- prune.tree(hospital_tree_all, best = 2)
plot(prune_hospital_train_all)
text(prune_hospital_train_all, pretty = 0)


```


Test set tree
```{r}
hospital_test_tree_all <- predict(hospital_tree_all, newdata = hospitals_test)

test_MSE <- mean((hospital_test_tree_all - hospitals_test$hospital_overall_rating)^2)

test_MSE

```



#REGRESSION TREE, removing patient rating as variable

```{r}
library(tree)

#Creating regression tree
hospital_tree_no_patient <- tree(hospital_overall_rating ~ region_num + hospital_type_num + hospital_ownership_num , data = hospitals_train)

plot(hospital_tree_no_patient)
text(hospital_tree_no_patient, pretty = 0)

```

Cross-validation of this tree
```{r cross-validation of the tree}
cv_hospital_tree_no_patient <- cv.tree(hospital_tree_no_patient)

plot(cv_hospital_tree_no_patient)
text(cv_hospital_tree_no_patient)

prune_hospital_train_no_patient <- prune.tree(hospital_tree_no_patient, best = 2)
plot(prune_hospital_train_no_patient)
text(prune_hospital_train_no_patient, pretty = 0)


```


Test set tree--no patient variable
```{r}
hospital_test_tree_no_patient <- predict(hospital_tree_no_patient, newdata = hospitals_test)

test_MSE <- mean((hospital_test_tree_no_patient - hospitals_test$hospital_overall_rating)^2)

test_MSE

```


Test MSE is higher than for original tree--should we even include this in our analysis? 


#RandomForest
```{r}
library(randomForest)
rf_hospitals <- randomForest(hospital_overall_rating ~ region_num + hospital_type_num + hospital_ownership_num + patient_experience_num , data = hospitals_train, importance = TRUE)

pred_rf_hospitals_test <- predict(rf_hospitals, newdata = hospitals_test)

varImpPlot(rf_hospitals)

rf_test_error <- mean((pred_rf_hospitals_test - hospitals_test$hospital_overall_rating)^2)

rf_test_error
```
